---
title: "Edit a project"
author: "Marion Shadbolt"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Edit_a_project}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ranglr)
library(magrittr)
```

Here I'll showcase useful functions when editing a projects

## Query the ingest API for uuids

This can be done using the function `query_existing_submission`. The user needs to provide the submission id for the submission they want to get uuids for. At the moment the easiest place to get the `submission_id` is from the staging and prod data trackers.

Prod data tracker: https://tracker.data.humancellatlas.org/
Staging data tracker: https://tracker.staging.data.humancellatlas.org/

The function returns a tibble converted from the API call JSON with many fields that can then be filtered to ones the user is interested in. The uuid itself is called `uuid.uuid`

```{r get_processes_uuids_staging}
staging_submission_id <- "5cdbe23d64198100078914ea"
processes_df <- query_existing_submission(staging_submission_id, 
                          submission_environment = "staging", 
                          entity_type = "processes")
tibble::glimpse(processes_df)
processes_df %>%
  dplyr::select(content.process_core.process_id, uuid.uuid)
```

This can be done for any of the 5 entity types, i.e. `biomaterials`, `files`, `processes`, `projects`, `protocols`.

```{r get_protocols_uuids_staging}
protocols_df <- query_existing_submission(staging_submission_id, 
                          submission_environment = "staging", 
                          entity_type = "protocols")
tibble::glimpse(protocols_df)
protocols_df %>%
  dplyr::select(content.protocol_core.protocol_id, content.protocol_core.protocol_name, 
                uuid.uuid)
```


And you can do the same in the prod ingest api

```{r get_uuids_prod}
prod_submission_id <- "5d3b11bc9be88c0008a9d744"

query_existing_submission(prod_submission_id, 
                          submission_environment = "prod", 
                          entity_type = "protocols")
```

Can get the `Ingest Submission id` from the prod data tracker https://tracker.data.humancellatlas.org/

